New Exchange Model for Child Pornography Analysis

### 2013-05-22 paper.tex
New paper!

\begin{document}
%
% --- Author Metadata here --- DO NOT REMOVE OR CHANGE 
\conferenceinfo{15$^{th}$ Twente Student Conference on IT}{June 21$^{st}$, 2012, Enschede, The Netherlands.}
\CopyrightYear{2012} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{New exchange model for child pornography analysis [abstract V1 | sec1 V9 | sec2 V7 | sec3 V7 | sec4 V5 | sec5 V2]}

@@

\maketitle
\begin{abstract}
This research concentrates around the topic of the criminal exchange of child pornography material on the Internet. The existing scientific material on this exchange lacks technical details that could help police detectives work more efficiently in child pornography cases. This research first takes a look at the existing literature on criminal child pornography exchange models. Thereafter it will look at the existing literature on new Internet based communication media, especially anonymization techniques. After this, a new criminal child pornography exchange model is presented and evaluated by interviewing police detectives concerned with online child pornography cases. The result of this research is a new criminal child pornography exchange model which allows for modeling anonymization within all the steps of the process. 
\end{abstract}

\keywords{child pornography, high tech crime, criminal exchange model}

\section{Introduction}
With the arrest of Robert M. in December 2010\cite{site:bbc} the Netherlands' police agency caught, allegedly, the biggest child abuser ever in the Netherlands. On his computer detectives found (hidden in a partially deleted TrueCrypt folder) 8 terabyte containing images and videos of child abuse. Along with the images and videos came the digital equivalent of 64.000 A4 papers containing chat logs with other child abusers, which led to the arrest of numerous of other child abusers in the Netherlands, Norway, Germany, the United States, Australia and many more countries. These arrests led to the question whether current detection methods used by police agencies to catch child abusers are still valid for the new internet era \cite{wijk2009achter}. The current detection methods are based upon exchange models of child pornography (CP) which may not really be adapted to this new era and therefore they may need to be revised. This research will try to answer the question of how to revise such exchange models to satisfy the need of police detectives for better analysis and detection methods.

Digital CP can be defined in multiple ways, this research will use a definition that is derived from the legal definition given by \cite{beech2008internet}, which cites the PROTECT Act of 2003:
\begin{quote}
	"...it is a crime to posses, manufacture or distribute pornography containing visual depictions of (a) a real child engaging in sexual acts or (b) a digital image, computer image, or computer generated image that is, or is indistinguishable from that of a minor engaged in sexually explicit conduct." from (18 U.S. Code Paragraph 2256(8) (B))
\end{quote}
This research acknowledges the legal definition of CP but it is not taking into account the issues related to computer generated images since they need to use different solutions than the ones discussed in the paper. Furthermore images and videos are widened to encompass digital material, leading to the following CP definition that is used by this research:
\begin{quote} 
	Digital child pornography is defined as digital pornography containing visual depictions of a real child engaging in sexual acts or a minor engaged in sexually explicit conduct. 
\end{quote}

To be able to describe the criminal process involved in the distribution, a type of model should be introduced to model the different steps involved in the process. This research will use Crime Schemes as proposed in \cite{willison2006understanding}. Steps in this process follow a flow. This means that certain processes depend on others, which need to be performed first. Groups of processes (categories) belong to each other and together form a category in the simplified model of the crime. In stealing a bike for example, there are different steps in how to break the lock, but they belong to each other and form the 'breaking the lock' category in the simplified model. Some different paths exist in breaking the lock, but it is always right to consider that one first needs to break the lock, before removing it from the bike. Since this research investigates exchange models and describe them via crime schemes, such an exchange model will from now on be called criminal exchange model.

@@

The main research question in this research is 'How can the currently used main criminal exchange models of child pornography be adapted to describe (1) the current child pornography user behavior and (2) the new Internet based communication media used by child pornography offenders?'. This main research question is divided into four different subquestions:
\begin{enumerate}
	\item What are the currently used main criminal exchange models of CP?
	\item What are the main differences between the communication media used in the past and today in the exchange of CP?
	\item How can the changes associated with the current CP user behavior and the new internet based communication media be reflected in the currently used main criminal exchange models of CP? 
\end{enumerate}

The research questions (1) and (2) are answered using a literature study. Finally, research question (3) is answered by proposing a new CP criminal exchange model and by verifying it using interviews with police detectives that are concerned with the recent arrests of several child pornography offenders.

This paper is organized as follows. In section 2 an overview of the current criminal exchange models of CP is given. Furthermore, this section provides the answer of the research question (1). Section 3 provides the technical enhancements on the CP criminal exchange models, providing the answer to research question (2). In Section 4 a new CP criminal exchange model is proposed that is verified by using interviews with police detectives that are concerned with the recent arrests of several child pornography offenders.. Moreover, this section answers research question (3). Section 5 concludes and provides recommendations for future work.

@@

Several papers and books briefly mention the exchange of CP, such as \cite{wolak2005child, jenkins2001beyond, webb2007characteristics}, which briefly focus on CP exchange via Internet, \cite{quayle2002paedophiles, beech2008internet}, which briefly focus on the CP exchange via usenet, bulletin boards or websites. However, only few of these papers that describe the CP exchange, 	go into more detail, see e.g. \cite{en2011phishing,wortley2006child,wijk2009achter,callanan2009internet}. The description level of detail used by papers is quite different and will therefore be elaborated in different sections.

@@

\begin{itemize}
	\item Completeness of the criminal exchange model: this refers to the number of categories modeled by the criminal exchange model. 
		The steps defined in this research that need to be modeled by a criminal exchange model are:
		\begin{itemize}
			\item \textit{Production}: The physical abuse and the recording of it. Also location selection, victim recruiting, editing, lighting, restructuring of the video, etc. fall in this category since this research is not focused upon this category and only needs a way to model all processes that are not involved in the criminal exchange model.
			\item \textit{Hiding}: Hiding information that could lead to the producer or distributor, by using techniques such as encryption and stenography \cite{krone2005international}.
			\item \textit{Distribution}: Making CP material available for others to download/view.
			\item \textit{Downloading}: Downloading or streaming CP material.
			\item \textit{Storage}: Storing CP material and possibly hiding it.
		\end{itemize}
		In table \ref{tables:exchangeModelsCompared}, the grade ++ represents the situation that all the steps mentioned above are supported, while - - means that none of these steps are supported.
	\item Modeling of Anonymity: this refers to whether the criminal exchange model, models in detail the support of anonymity in the criminal CP process. Anonymity support refers to support of activities in such a way that the identity of the CP user cannot be detected. In table \ref{tables:exchangeModelsCompared}, the grade ++ represents the situation that a criminal exchange model provides all means to model Anonymity support. The grade - - represents the situation that the criminal exchange model does not  provide any means of modeling Anonymity support.
	\item Modeling of Distribution types: this refers to whether the criminal exchange model, models in detail the support of Distribution type in the CP process. Distribution type refers to the way of how the CP media is distributed. In table \ref{tables:exchangeModelsCompared}, the grade ++ represents the situation that a criminal exchange model provides all means to model the Distribution type. The grade - - represents the situation that the criminal exchange model does not provide any means of modeling the Distribution type.
	\item Transaction type: this refers to the type of transactions involved within the Criminal CP process. These transactions can be either Financial means (money) or CP interchange material. 
\end{itemize}

\subsection{Faber et al., 2011 \cite{en2011phishing}}
Figure \ref{fig:faber_01} gives the visual representation of the criminal CP exchange model proposed by \cite{en2011phishing}. This figure shows four different categories (\textit{Production},\textit{Hiding}, \textit{Distribution} and \textit{Storage}). \cite{en2011phishing} groups different steps in the process of the CP exchange. The step \textit{editing} is not included in such a group since it is not considered as being an important step in the criminal CP scheme.

Faber et al. \cite{en2011phishing} is the only research document that describes some sort of criminal exchange model for the criminal process of CP. This criminal CP exchange model \cite{en2011phishing} describes the process that is sketched in Figure \ref{fig:faber_01}. The \textit{Editing} event is just a video editing event and therefore it can be left out of the exchange model. The different types of steps that could be taken in the \textit{Coding} event are similar to what other authors have concluded (i.e. \cite{wortley2006child, jenkins2001beyond,wolak2005child}). It includes untraceability, invisibility, illegibility and unrecognizability. The \textit{Payment} event does not only involve cash payments, but also includes interchange of CP material which is found in several other articles as well (i.e. \cite{wortley2006child,wijk2009achter,wolak2005child}). The \textit{Distribution} event only covers the describing of different communication media such as newsgroups \cite{site:usenet}, bulletin boards \cite{site:bulletin}, P2P \cite{site:p2p}, e-mail, IRC \cite{site:IRC}, FTP \cite{site:FTP} and websites, but lacks to clarify the specifics for CP exchange on those communication media or the amount of usage on these different communication media. The authors of \cite{en2011phishing} make a clear distinction between the exchange via a website (acquire domain, acquire hosting and distribution) and exchange via other media (acquire costumers, payment, distribution). This is in clear contrast with different other papers, \cite{wortley2006child,wijk2009achter,callanan2009internet}, that all state that using a website to distribute CP always includes a payment.

@@

The \textit{abuse recording} step is thoroughly described, covering topics as the severity of the recordings \cite{quayle2002paedophiles}, but it also covers topics such as the victim selection, grooming, consolidation, staff recruitment and location selection. The \textit{editing} step involves the anonymization of the abuse, making scenes, compiling it, making descriptions, etc. The \textit{acquire domain} and \textit{hosting} involve the process of getting a website and domain name to spread the material. This can be bought online using (stolen) credit cards, web money or servers that could be hacked. The \textit{acquire costumers} event describes the process of getting customers to find the CP material, in addition to websites, it involves mass mailing and it provides adds on the normal porn websites or getting inside costumers from fora or networks. The \textit{payment} event could be either money (using (stolen) credit cards, web money, billing companies) or interchange of CP material. The \textit{distribution} process involves the uploading and downloading of the CP material via virtual communities, news groups, fora, p2p, e-mail, chat, FTP and websites. The \textit{storage} step involves storing the material and hiding it from authorities and other users of that computer or network.

The processes modeled by Faber et al. involving the abuse and pre-abuse steps are in this research modeled to one process called \textit{abuse recording}, and therefore it is classified into the \textit{Production} category. The steps modeled in the process \textit{coding} are used to hide the producer's and distributor's identity and is therefore classified into the \textit{Hiding} category. The processes that cover the \textit{acquire costumers} steps and the steps of leading them to the CP material are grouped together with the actual distribution and payment processes to form the \textit{Distribution} category. The storage process modeled by Faber et al. is considered to be the \textit{Storage} category in this research.

Faber et al. elaborates thoroughly on the criminal CP exchange model and therefore are rated with a + on the completeness of the criminal model. The model lacks on providing details on modeling the anonymity and the ability to model Anonymity is therefore rated with a -. The up-to-date distribution media also lack and is therefore rated with a -. The types of transactions involved are both money and the interchange of CP material, see Table \ref{tables:exchangeModelsCompared}.

@@

The \textit{production} process models the production of material, mostly made by amateurs but also covers professional recordings. The \textit{distribution} process describes in short the existence of pedophile rings or organized crime groups that interchange the CP material on different platforms like websites, e-mail, instant messages, newsgroups, bulletin boards, chat and p2p. The \textit{downloading} process describes the actual search for CP material by CP users. In \cite{wortley2006child} is stated that mass mailings and the ads on normal porn websites hardly exist anymore.

The process described by Wortley et al. as the \textit{production} process is modeled as the \textit{Production} category. The \textit{distribution} process is modeled as the \textit{Distribution} category and the \textit{downloading} process is modeled as the \textit{Downloading} category.

Wortley et al. do not elaborate thoroughly on the criminal CP exchange model, therefore the completeness of the criminal model is rated with a + / -. The modeling of Anonymity is not supported and therefore rated with a - -. The ability to model up-to-date distribution media is a little bit better and therefore rated with a -.  The type of transactions involved is only the interchange of CP material, see Table \ref{tables:exchangeModelsCompared}.

@@

Van Wijk et al. \cite{wijk2009achter} describes four different types of distribution; commercial distribution (via mail delivery), commercial websites (pay by credit card or emoney), open networks (open IRC channels, open P2P networks, open bulletin boards) and closed networks (chat, e-mail, closed bulletin boards and closed P2P networks). Because of recent police activity against CP, more or less all distribution is in the last category, see \cite{wortley2006child, callanan2009internet}. Van Wijk et al. takes a closer look into the high tech components involved in the \textit{distribution} process. However, the focus is more on the CP downloaders instead of being on the CP distributors. Some of the mentioned high tech components involve (1) virtual hard drives to browse without leaving traces on the computer \cite{site:VHD}, (2) the use of disk encryption software to hide downloaded material on specific parts of the hard drive \cite{site:diskenc}, (3) the use of proxy servers as storage locations or to browse anonymously \cite{site:proxy} and (4) software to erase a hard drive.

@@

The process described by Van Wijk et al. as \textit{production} is mentioned in the model, but no further information is given about this step. The \textit{distribution} step is different from earlier models and elaborates more on the different distribution media. The distribution is split in four different types of distribution.
\begin{itemize}
	\item Commercial distribution: uses post-order to order CP material online and receive it by post mail. 
	\item Commercial websites: websites that run on servers mostly located in eastern European cities, which are distributing the CP material by selling it online.
	\item Open networks: networks, which can be used to distribute CP material without requiring any password or security protection. Such open networks are: open P2P networks, open chat rooms and open bulletin boards.
	\item Closed networks: networks that can be used to distribute CP material using the same media as the open networks, but which are closed by using password and security protection, such as encryption and decryption.
\end{itemize}

The \textit{production} process is modeled as the \textit{Production} category. The different distribution media are grouped together as the \textit{Distribution} category since all these distribution media types involve the characteristics of this category. The \textit{downloading} process is modeled as the \textit{Downloading} category and the \textit{storage} process is modeled as the \textit{Storage} category.

Van Wijk et al. elaborates on the criminal exchange model and the completeness of the criminal model is therefore rated with a+. It also presents some details on the Anonymity support and that is therefore rated a +. The ability to model up-to-date distribution media is relatively extensive and therefore rated with a +. The types of transactions involved are encompassing both money and interchange of CP material, see Table \ref{tables:exchangeModelsCompared}.

@@

Callanan et al. \cite{callanan2009internet} describes different communication media used to distribute CP material than the other CP exchange models described in the previous subsections. In particular, \cite{callanan2009internet} describes Freenet \cite{site:freenet} as a way to distribute CP material in encrypted and sliced packets which are then on a P2P manner distributed to users of Freenet (including non CP users). Freenet supports multiple anonymity options, making it possible to easily hide the identity of its users. Furthermore, \cite{callanan2009internet} describes several characteristics of Usenet system and the incentives of why this system is used among CP users. Both these communication media (i.e., \textit{distribution} process) have the characteristic that it is really hard to delete the CP material once it has been distributed.

@@

The hiding identity process describes different communication media that could be used either by downloaders of distributors for the exchange of CP. The actual downloading process models anonymity support by those communication media. The distribution process describes the anonymity support offered to distributors of CP material.

\cite{callanan2009internet} describes various methods of identity hiding, which in turn could be used for downloading and/or distribution. The \textit{hiding} identity process describes different communication media that could be used by either CP material downloaders or CP material distributors. The actual \textit{downloading} process encompasses the anonymity support used by those communication media. The \textit{distribution} process encompasses the anonymity support offered to distributors of CP material. The \textit{hiding identity} process as modeled by Callanan et al. cannot be classified on any of the categories introduced at the introductory part of this section. In particular, this process encompasses both the identity hiding of CP material downloaders and CP material distributors. The \textit{downloading} process is modeled as the \textit{Downloading} category and the \textit{distribution} process is modeled as the \textit{Distribution} category.

Callanan et al. does not elaborate on the criminal exchange model and the completeness of the criminal model is therefore rated a -. The detailed description of Anonymity support is rated with a + +. The detailed description of up-to-date distribution media is also rated with a + +, since it is very elaborative. The types of transactions involved are not discussed, see Table \ref{tables:exchangeModelsCompared}.

\subsection{Conclusion}
A comparison between the criminal CP exchange models described in this section is depicted in Table \ref{tables:exchangeModelsCompared}. The criteria used for this comparison are given in the introductory part of section 2.1.
From this table it can be seen that the only detailed criminal CP exchange model is found in only \cite{en2011phishing}. The components by this model are also mentioned by other authors as well, except for the lack of payment in distribution via websites, which is not covered in \cite{en2011phishing}. The amount of usage of different communication media has made a switch from open networks to more closed ones as found by \cite{wortley2006child,wijk2009achter}. The types of CP payment methods mentioned in the exchange models differ. Though, it can be stated that the more advanced a CP user is, the less likely he/she will favorit paying for CP material. In particular, he/she will be far more interested in interchanging CP material, and thereby becoming more and more important in the online world of CP exchange \cite{wijk2009achter,quayle2002paedophiles,krone2005international,en2011phishing}. \cite{wijk2009achter} describes several high tech components involved in the CP material \textit{downloading} process and its \texttt{distribution} (virtual hard disks, disk encryption software and proxy servers). \cite{callanan2009internet} describes in more detail several communication media types (Usenet and Freenet) and the anonymity support used for the actual CP material distribution. However, \cite{callanan2009internet} is not providing any information on the used transaction types. \cite{en2011phishing} and \cite{wijk2009achter} use both Money and Interchange of CP material as transaction types, while \cite{wortley2006child} uses only the interchange of CP material as transaction type.

@@

\section{Enhancements of CP exchange model components}
This section describes the ongoing technical enhancements on the CP criminal exchange models. The technical enhancements that will be discussed in this section will also be rated on the scale that the different criminal exchange models were rated on in section 2. The results can be found in Table \ref{tables:dataExchangeCompared}.

>>>>

\begin{document}
%
% --- Author Metadata here --- DO NOT REMOVE OR CHANGE 
\conferenceinfo{17$^{th}$ Twente Student Conference on IT}{June 25$^{th}$, 2012, Enschede, The Netherlands.}
\CopyrightYear{2012} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

\title{New Exchange Model for Child Pornography Analysis}

@@

\maketitle
\begin{abstract}
This research concentrates around the topic of the criminal exchange of child pornography material on the Internet. The existing scientific material on this criminal exchange lacks technical details that could help police detectives work more efficiently in child pornography cases. This paper specifies a new criminal exchange model of child pornography that describes (1) the current child pornography user behavior and (2) the new Internet based communication media and anonymization techniques used by child pornography offenders. The specified criminal exchange model has been evaluated and verified by interviewing police detectives concerned with online child \\pornography cases.
\end{abstract}

\keywords{child pornography, high tech crime, criminal exchange\\ model}

\section{Introduction}
With the arrest of Robert M. in December 2010 \cite{site:bbc} the Netherlands' police agency caught, allegedly, the biggest child abuser ever in the Netherlands. On his computer, detectives found (hidden in a partially deleted TrueCrypt folder) eight terabyte containing images and videos of child abuse. Along with the images and videos came the digital equivalent of 64.000 A4 papers containing chat logs with other child abusers, which led to the arrest of numerous of other child abusers in the Netherlands, Norway, Germany, the United States, Australia and many more countries. These arrests led to the question whether current detection methods used by police agencies to catch child abusers are still valid for the new Internet era \cite{wijk2009achter}. The current detection methods are based upon exchange models of child pornography (CP) which may not really be adapted to this new era and therefore they may need to be revised. This research will try to answer the question of how to revise such exchange models to satisfy the need of police detectives for better analysis and detection methods.

Digital CP can be defined in multiple ways, this research will use a definition that is derived from the legal definition given by \cite{beech2008internet}, which cites the PROTECT Act of 2003:
\begin{quote}
	"...it is a crime to posses, manufacture or distribute pornography containing visual depictions of (a) a real child engaging in sexual acts or (b) a digital image, computer image, or computer generated image that is, or is indistinguishable from that of a minor engaged in sexually explicit conduct." from (18 U.S. Code Paragraph 2256(8) (B))
\end{quote}
This research acknowledges the legal definition of CP but it is not taking into account the issues related to computer generated images since they need to use different solutions than the ones discussed in the paper. Furthermore, images and videos are widened to encompass digital material, leading to the following CP definition that is used by this research:
\begin{quote} 
	Digital child pornography is defined as digital pornography containing visual depictions of a real child engaging in sexual acts or a minor engaged in sexually explicit conduct. 
\end{quote}

In order to be able to describe the criminal process involved in the distribution, a type of model should be introduced to model the different steps involved in the process. This research will use Crime Schemes as proposed in \cite{willison2006understanding}. Steps in this process follow a flow. This means that certain processes depend on others, which need to be performed first. Groups of processes (categories) belong to each other and together form a category in the simplified model of the crime. In stealing a bike for example, there are different steps in how to break the lock, but they relate to each other and form the 'breaking the lock' category in the simplified model. Some different paths exist in breaking the lock, but it is always right to consider that one needs to first break the lock, before removing it from the bike. Since this research investigates exchange models and describes them via crime schemes, such an exchange model will from now on be called a criminal exchange model.

@@

The main research question in this research is `How can the currently used main criminal exchange models of child pornography be adapted to describe (1) the current child pornography user behavior and (2) the new Internet based communication media and anonymization techniques used by child pornography offenders?'. This main research question is divided into three different sub questions:
\begin{enumerate}
	\item What are the currently used main criminal exchange models of CP?
	\item What are the main differences between the communication media used in the past and today in the exchange of CP?
	\item How can the changes associated with the current CP user behavior and the new internet based communication media be reflected in the currently used main criminal exchange models of CP? 
\end{enumerate}

The research questions (1) and (2) are answered using a literature study. Finally, research question (3) is answered by proposing a new CP criminal exchange model and by verifying it using interviews with seven police detectives that are concerned with the recent arrests of several child pornography offenders.

This paper is organized as follows. In Section 2 an overview of the current criminal exchange models of CP is given. Furthermore, this section provides the answer of the research question (1). Section 3 provides the technical enhancements on the CP criminal exchange models, providing the answer to research question (2). In Section 4 a new CP criminal exchange model is proposed, which is verified by interviewing police detectives that are involved with the recent arrests of several child pornography offenders. Moreover, this section answers research question (3). Section 5 concludes and provides recommendations for future work.

@@

Several papers and books briefly mention the exchange of CP, such as \cite{wolak2005child, jenkins2001beyond, webb2007characteristics}, which briefly focus on CP exchange via Internet and, \cite{quayle2002paedophiles, beech2008internet}, which briefly focus on the CP exchange via usenet, bulletin boards or websites. However, only few of these papers that describe the CP exchange, 	go into more detail, see e.g. \cite{en2011phishing,wortley2006child,wijk2009achter,callanan2009internet}. The description level of detail used by papers is quite different and will therefore be elaborated in different sections.

@@

\begin{itemize}
	\item Completeness of the criminal exchange model: this refers to the number of categories modeled by the criminal exchange model. 
		The steps defined in this research, which need to be modeled by a criminal exchange model are:
		\begin{itemize}
			\item \textit{Production}: The physical abuse and the recording of it. Also location selection, victim recruiting, editing, lighting, restructuring of the video, etc. fall in this category, since this research is not focused upon this category and only needs a way to model all processes that are not involved in the criminal exchange model.
			\item \textit{Hiding}: Hiding information that could lead to the producer or distributor, by using techniques such as encryption and stenography \cite{krone2005international}.
			\item \textit{Distribution}: Making CP material available for others to download/view.
			\item \textit{Downloading}: Downloading or streaming CP material.
			\item \textit{Storage}: Storing CP material and possibly hiding it.
		\end{itemize}
		In Table \ref{tables:exchangeModelsCompared}, the grade ++ represents the situation that all the steps mentioned above are supported, while - - means that none of these steps are supported.
	\item Modeling of Anonymity: this refers to whether the criminal exchange model, models in detail the support of anonymity in the criminal CP process. \\Anonymity support refers to the support of activities in such a way that the identity of the CP user cannot be detected. In Table \ref{tables:exchangeModelsCompared}, the grade ++ represents the situation that a criminal exchange model provides all means to model Anonymity support. The grade - - represents the situation that the criminal exchange model does not  provide any means of modeling Anonymity support.
	\item Modeling of Distribution types: this refers to whether the criminal exchange model, models in detail the support of Distribution type in the CP process. Distribution type refers to the way of how the CP media is distributed. In Table \ref{tables:exchangeModelsCompared}, the grade ++ represents the situation that a criminal exchange model provides all means to model the Distribution type. The grade - - represents the situation that the criminal exchange model does not provide any means of modeling the Distribution type.
	\item Transaction type: this refers to the type of transactions involved within the Criminal CP process. These transactions can be either Financial means (money) or CP interchange material. 
\end{itemize}

\subsection{Faber et al., 2011 \cite{en2011phishing}}
Figure \ref{fig:faber_01} gives the visual representation of the criminal CP exchange model proposed by \cite{en2011phishing}. This figure shows four different categories (\textit{Production},\textit{Hiding}, \textit{Distribution} and \textit{Storage}). \cite{en2011phishing} groups different steps in the process of the CP exchange. The step \textit{Editing} is not included in such a group since it is not considered as being an important step in the criminal CP scheme.

Faber et al. \cite{en2011phishing} is the only research document that describes some sort of criminal exchange model for the criminal process of CP. This criminal CP exchange model \cite{en2011phishing} describes the process that is sketched in Figure \ref{fig:faber_01}. The \textit{Editing} event is just a video editing event and therefore it can be left out of the exchange model. The different types of steps that could be taken in the \textit{Coding} event are similar to what other authors have concluded (i.e. \cite{wortley2006child, jenkins2001beyond,wolak2005child}). It includes untraceability, invisibility, illegibility and unrecognizability. The \textit{Payment} event does not only involve cash payments, but also includes interchange of CP material that is found in several other articles as well (i.e. \cite{wortley2006child,wijk2009achter,wolak2005child}). The \textit{Distribution} event only covers the description different communication media such as newsgroups \cite{site:usenet}, bulletin boards \cite{site:bulletin}, P2P \cite{site:p2p}, e-mail, IRC \cite{site:IRC}, FTP \cite{site:FTP} and websites, but lacks to clarify the specifics for CP exchange on those communication media or the amount of usage on these different communication media. The authors of \cite{en2011phishing} make a clear distinction between the exchange via a website (acquire domain, acquire hosting and distribution) and exchange via other media (acquire costumers, payment, distribution). This is in clear contrast with different other papers, \cite{wortley2006child,wijk2009achter,callanan2009internet}, which all state that using a website to distribute CP always includes a payment.

@@

The \textit{Abuse Recording} step is thoroughly described, covering topics as the severity of the recordings \cite{quayle2002paedophiles}, but it also covers topics such as the victim selection, grooming, consolidation, staff recruitment and location selection. The \textit{Editing} step involves the anonymization of the abuse, making scenes, compiling it, making descriptions, etc. The \textit{Acquire Domain} and \textit{Hosting} involve the process of getting a website and domain name to spread the material. This can be bought online using (stolen) credit cards, web money or servers that could be hacked. The \textit{Acquire Customers} event, describes the process of getting customers to find the CP material, in addition to websites, it involves mass mailing and it provides adds on the normal porn websites or getting inside costumers from fora or networks. The \textit{Payment} event could be either money (using (stolen) credit cards, web money, billing companies) or interchange of CP material. The \textit{Distribution} process involves the uploading and downloading of the CP material via virtual communities, news groups, fora, p2p, e-mail, chat, FTP and websites. The \textit{Storage} step involves storing the material and hiding it from authorities and other users of that computer or network.

The processes modeled by Faber et al. involving the abuse and pre-abuse steps are in this research modeled to one process called \textit{Abuse Recording}, and therefore it is classified into the \textit{Production} category. The steps modeled in the process \textit{Coding} are used to hide the producer's and distributor's identity and is therefore classified into the \textit{Hiding} category. The processes that cover the \textit{Acquire Customers} steps and the steps of leading them to the CP material are grouped together with the actual distribution and payment processes to form the \textit{Distribution} category. The storage process modeled by Faber et al. is considered to be the \textit{Storage} category in this research.

Faber et al. elaborates thoroughly on the criminal CP exchange model and therefore are rated with a + on the completeness of the criminal model. The model lacks on providing details on modeling the anonymity and the ability to model Anonymity Support is therefore rated with a -. Faber et al. also lacks on providing up-to-date distribution media and is therefore rated with a -. The types of transactions involved are both Money and the Interchange of CP material, see Table \ref{tables:exchangeModelsCompared}.

@@

The \textit{Production} process models the production of material, mostly made by amateurs but also covers professional recordings. The \textit{Distribution} process describes in short the existence of pedophile rings or organized crime groups that interchange the CP material on different platforms like websites, e-mail, instant messages, newsgroups, bulletin boards, chat and p2p. The \textit{Downloading} process describes the actual search for CP material by CP users. In \cite{wortley2006child}, it is stated that mass mailings and the ads on normal porn websites hardly exist anymore.

The process described by Wortley et al. as the \textit{Production} process is modeled as the \textit{Production} category. The \textit{Distribution} process is modeled as the \textit{Distribution} category and the \textit{Downloading} process is modeled as the \textit{Downloading} category.

Wortley et al. do not elaborate thoroughly on the criminal CP exchange model, therefore the completeness of the criminal model is rated with a + / -. The modeling of Anonymity Support is not supported and therefore rated with a - -. The ability to model up-to-date distribution media is a little bit better and therefore rated with a -.  The type of transactions involved is only the Interchange of CP material, see Table \ref{tables:exchangeModelsCompared}.

@@

Van Wijk et al. \cite{wijk2009achter} describes four different types of distribution; commercial distribution (via mail delivery), commercial websites (pay by credit card or emoney), open networks (open IRC channels, open P2P networks, open bulletin boards) and closed networks (chat, e-mail, closed bulletin boards and closed P2P networks). Due to recent police activity against CP, more or less all distribution belongs to the last category, i.e. closed networks, see \cite{wortley2006child, callanan2009internet}. Van Wijk et al. takes a closer look into the high tech components involved in the \textit{distribution} process. However, the focus is more on the CP downloaders instead of being on the CP distributors. Some of the mentioned high tech components involve (1) virtual hard drives to browse without leaving traces on the computer \cite{site:VHD}, (2) the use of disk encryption software to hide downloaded material on specific parts of the hard drive \cite{site:diskenc}, (3) the use of proxy servers as storage locations or to browse anonymously \cite{site:proxy} and (4) software to erase a hard drive.

@@

The process described by Van Wijk et al. as \textit{Production} is mentioned in the model, but no further information is given about this step. The \textit{Distribution} step is different from earlier models and elaborates more on the different distribution media. The distribution is split in four different types of distribution.
\begin{itemize}
	\item Commercial distribution: uses post-order to order CP material online and receive it by post mail. 
	\item Commercial websites: websites that run on servers mostly located in eastern European cities, which are distributing the CP material by selling it online.
	\item Open networks: networks, which can be used to distribute CP material without requiring any password or security protection. Such open networks are: open P2P networks, open chat rooms and open bulletin boards.
	\item Closed networks: networks that can be used to distribute CP material using the same media as the open networks, but which are closed by using password and security protection, such as encryption and decryption.
\end{itemize}

The \textit{Production} process is modeled as the \textit{Production} category. The different distribution media are grouped together as the \textit{Distribution} category since all these distribution media types involve the characteristics of this category. The \textit{Downloading} process is modeled as the \textit{Downloading} category and the \textit{Storage} process is modeled as the \textit{Storage} category.

Van Wijk et al. elaborates on the criminal exchange model and the completeness of the criminal model is therefore rated with a +. It also presents some details on the Anonymity support and therefore it is rated with a +. The ability to model up-to-date distribution media is relatively extensive and therefore rated with a +. The types of transactions involved are encompassing both Money and Interchange of CP material, see Table \ref{tables:exchangeModelsCompared}.

@@

Callanan et al. \cite{callanan2009internet} describes different communication media used to distribute CP material than the other CP exchange models described in the previous subsections. In particular, \cite{callanan2009internet} describes Freenet \cite{site:freenet} as a way to distribute CP material in encrypted and sliced packets which are then distributed to users of Freenet on a P2P manner (including non CP users). Freenet supports multiple anonymization options, making it possible to easily hide the identity of its users. Furthermore, \cite{callanan2009internet} describes several characteristics of Usenet system and the incentives of why this system is used among CP users. Both these communication media (i.e., \textit{Distribution} process) have the characteristic that it is really hard to delete the CP material once it has been distributed.

@@

The \textit{Hiding Identity} process describes different communication media that could be used either by downloaders of distributors for the exchange of CP. The actual downloading process models Anonymity Support by those communication media. The distribution process describes the Anonymity Support offered to distributors of CP material.

\cite{callanan2009internet} describes various methods of identity hiding, which in turn could be used for downloading and/or distribution. The \textit{Hiding} identity process describes different communication media that could be used by either CP material downloaders or CP material distributors. The actual \textit{Downloading} process encompasses the anonymity support used by those communication media. The \textit{Distribution} process encompasses the anonymity support offered to distributors of CP material. The \textit{Hiding Identity} process as modeled by Callanan et al. cannot be classified in any of the categories introduced at the introductory part of this section. In particular, this process encompasses both the identity hiding of CP material downloaders and CP material distributors. The \textit{Downloading} process is modeled as the \textit{Downloading} category and the \textit{Distribution} process is modeled as the \textit{Distribution} category.

Callanan et al. does not elaborate on the criminal exchange model and the completeness of the criminal model is therefore rated with a -. The detailed description of Anonymity support is rated with a + +. The detailed description of up-to-date distribution media is also rated with a + +, since it is very elaborative. The types of transactions involved are not discussed, see Table \ref{tables:exchangeModelsCompared}.

\subsection{Conclusion}
A comparison between the criminal CP exchange models described in this section is depicted in Table \ref{tables:exchangeModelsCompared}. The criteria used for this comparison are given in the introductory part of section 2.1.
From this table it can be seen that the only detailed criminal CP exchange model is found in \cite{en2011phishing}. The components of this model are mentioned by other authors as well, except for the lack of payment in distribution via websites, which is not covered in \cite{en2011phishing}. The amount of usage of different communication media has made a switch from open networks to more closed ones as found by \cite{wortley2006child,wijk2009achter}. The types of CP payment methods mentioned in the exchange models differ. Though, it can be stated that the more advanced a CP user is, the less likely he/she will favor paying for CP material. In particular, he/she will be far more interested in interchanging CP material, and thereby becoming more and more important in the online world of CP exchange \cite{wijk2009achter,quayle2002paedophiles,krone2005international,en2011phishing}. \cite{wijk2009achter} describes several high tech components involved in the CP material \textit{Downloading} process and its \textit{Distribution} (virtual hard disks, disk encryption software and proxy servers). \cite{callanan2009internet} describes in more detail several communication media types (Usenet and Freenet) and the anonymity support used for the actual CP material distribution. However, \cite{callanan2009internet} is not providing any information on the used transaction types. \cite{en2011phishing} and \cite{wijk2009achter} use both Money and Interchange of CP material as transaction types, while \cite{wortley2006child} uses only the interchange of CP material as transaction type.

@@

\section{Enhancements of CP exchange model components}
This section describes the ongoing technical enhancements on the CP criminal exchange models. The technical enhancements that will be discussed in this section will also be rated on the scale that the different criminal exchange models were rated on in Section 2. The results can be found in Table \ref{tables:dataExchangeCompared}.
